---
title: "Analytical Plan for Sensitivity of mortality rates to the imputation of missing socioeconomic data: cohort study"
subtitle: 'DOCUMENT: SAP-2023-017-BH-v01'
author: '**From:** Felipe Figueiredo **To:** Brennan Hickson'
date: "2023-04-25"
output:
  html_document:
    fig_caption: yes
    fig_height: 6
    fig_width: 6
    keep_md: yes
    number_sections: yes
    toc: yes
  pdf_document:
    number_sections: yes
    toc: yes
  word_document:
    fig_caption: yes
    fig_height: 6
    fig_width: 6
    reference_docx: misc/style_SAP_en.docx
    toc: yes
# toc-title: "Sumário"
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE)
knitr::opts_knit$set(root.dir = normalizePath(".."))
options(scipen = 999)
library(pander)
library(knitr)
panderOptions('table.style', 'rmarkdown')
panderOptions('table.split.table', 160)
```

---

\newpage

**Analytical Plan for Sensitivity of mortality rates to the imputation of missing socioeconomic data: cohort study**

**Document version**

```{r history}
Version <- c("01")
Changes <- c("Initial version")
history <- cbind(Version, Changes)
colnames(history) <- c("Version", "Alterations")
# pander(history, split.cells = 70)
kable(history)
```

```{r cleaning, include=FALSE}
source("scripts/input.R", encoding = 'UTF-8')
```

---

# Abbreviations

- FIM: Functional Independence Measure
- CI: confidence interval
- DCI: Distress community index
- HR: hazards ratio
- LOCF: Last observation carried forward
- NOCB: Next observation carried backward
- SD: standard deviation
- SES: socioeconomic status
- TBI: Traumatic brain injury

# Context

## Objectives

1. To describe the missingness in zip codes at each follow up collection;
1. To impute missing Zip codes with data available in previous follow up collections.
1. To assess the sensitivity of the association between mortality and socioeconomic status to the imputation of participant missing location.

## Hypotheses

Imputing the missing zip codes will decrease the missingness in the dataset and improve the model fit.

# Data

## Raw data

The original data base had
`r Nvar_orig`
variables collected on
`r Nobs_orig`
observations.

<!-- Upon study start the raw data will be collected in a raw table, that will be processed before analysis. -->
<!-- The raw dataset to be collected will have -->
<!-- `r Nvar_orig` -->
<!-- variables collected on -->
<!-- `r Nobs_orig` -->
<!-- observations. -->

<!-- This dataset will include the dates of entry and exit of the cohort, or the date of hospital admission and the date where the endpoint was reached (either an event, or hospital discharge). -->
<!-- Table 1 shows the structure of the raw dataset. -->

<!-- ```{r analytical data} -->
<!-- analytical_mockup %>% -->
<!--   pander(caption = "**Table 1** Raw dataset structure.") -->
<!-- ``` -->

<!-- Each row represents all information collected from each study participant, and each participant included will require a unique study ID. -->
<!-- The outcome should be recorded as a binary variable: either the study participant reached the endpoint -->
<!-- ([outcome]) -->
<!-- or reached the end of study period without experiencing the event. -->
<!-- This information can be recorded in either text form (eg, yes/no), or an indicator -->
<!-- ([outcome] = 1, -->
<!-- end of observation time or discharge = 0). -->
<!-- Exposure and all comorbidities should be recorded as binary variables, whenever possible. -->
<!-- All comorbidities should be recorded as binary variables, whenever possible. -->

<!-- Surgery length should be recorded in -->
<!-- minutes, -->
<!-- hours, -->
<!-- and ICU length in -->
<!-- days. -->
<!-- weeks. -->
<!-- months. -->

<!-- Surgery type will be recorded as a categorical variable. -->
<!-- It is recommended that the number of classes (types of surgery) be as small as possible, to avoid loss of precision in the analysis estimates, without the need to increase the sample size. -->
<!-- Broader categories of surgeries should be preferred, whenever possible. -->

## Analytical dataset

<!-- After the cleaning process -->
<!-- `r Nvar_final` -->
<!-- variables were included in the analysis -->
<!-- with -->
<!-- `r Nobs_final` -->
<!-- observations. -->
After the cleaning process
`r Nvar_final`
variables were included in the analysis.
The total number of observations excluded due to incompleteness and exclusion criteria will be reported in the analysis.
Table 1 shows the structure of the analytical dataset.

```{r analytical data}
analytical_mockup %>%
  pander(caption = "**Table 1** Analytical dataset structure after variable selection and cleaning.", split.table = Inf)
```

All variables in the analytical set were labeled according to the raw data provided and values were labeled according to the data dictionary for the preparation of production-quality results tables and figures.

# Study parameters

## Study design

This is a retrospective analysis of a prospective cohort study.

## Inclusion and exclusion criteria

**Inclusion criteria**

- Participants with at most 10 years of follow up;
- Participants included in the cohort between 2010-01-01 and 2018-12-31.

**Exclusion criteria**

Observations after 2019-12-31 will be excluded in order to mitigate risk of confounding by COVID-19 related deaths.
Observations prior to this date will still be considered for participants where such data is available.

## Exposures

SES of the neighborhood to which the participant was discharged.
The SES measure was stratified into its quintiles, and labelled according to the data dictionary to facilitate interpretation of the results.

## Outcomes

**Specification of outcome measures** (Zarin, 2011):

1. (Domain) Mortality
2. (Specific measurement) Death
3. (Specific metric) Time-to-event
4. (Method of aggregation) Hazard ratio

**Primary outcome**

Death after a brain injury.

## Covariates

- Sex
- Race
- Age at injury
- Substance Problem Use
- Education
- Employment status
- Rural area
- Previous seizure disorder diagnosis
- Spinal cord injury
- Cause of injury
- Primary rehabilitation payer
- Residence after rehab discharge
- Days From Injury to Rehab Discharge
- FIM Motor at Discharge
- FIM Cognitive at Discharge

# Statistical methods

## Statistical analyses

### Descriptive analyses

The epidemiological profile of the study participants will be described.
Demographic
and clinical variables
will be described as
mean (SD)
or as counts and proportions (%), as appropriate.
The distributions of participants' characteristics will be summarized in tables and visualized in exploratory plots.

### Inferential analyses

All inferential analyses will be performed in the statistical models (described in the next section).

### Statistical modeling

This analysis will evaluate the sensitivity to the model specification chosen in **SAR-2023-016-BH** to changes in the SES data (defined as the exposure of that analysis).
The model specification used for the sensitivity analysis will be the best model selected in that associated report.

For reference, the specification defined there regresses the hazard on the SES controlling for all covariates listed in section 4.5, except "previous seizure".

<!-- $$HR \sim exposure + Sex + Race + AGE + EDUCATION + -->
<!--   EMPLOYMENT + ResDis + RURALdc + RehabPay1 + SCI + PROBLEMUse + -->
<!--   FIMMOTD4 + FIMCOGD4$$ -->

This model specification will be applied on all datasets created from the imputation approaches described in section 5.1.4, and the Schoenfeld test will be applied to verify the proportional hazards assumption on all model terms.

### Missing data

A couple of simple imputation approaches will be applied on missing values for Zip codes, before the DCI data is merged into the TBI database.
An LOCF-based imputation will be applied to impute future Zip codes based on the last known value for each individual.
An additional dataset will be created by applying both NOCB- and LOCF-based imputations on missing values, with the intention of increasing the proportion of location data before the DCI data is merged and inclusion/exclusion criteria are applied, in particular the criterion that selects only the last valid observation of the individual in the study period.

## Significance and Confidence Intervals

All analyses will be performed using the significance level of 5%.
All significance hypothesis tests and confidence intervals computed will be
two-tailed.
<!-- left-tailed. -->
<!-- right-tailed. -->

## Study size and Power

N/A

## Statistical packages

This analysis will be performed using statistical software `R` version `r getRversion()`.

# Observations and limitations

**Recommended reporting guideline**

The adoption of the EQUATOR network (<http://www.equator-network.org/>) reporting guidelines have seen increasing adoption by scientific journals.
All observational studies are recommended to be reported following the STROBE guideline (von Elm et al, 2014).
<!-- All clinical trials are recommended to be reported following the CONSORT guideline (Schulz K F, Altman D G, Moher D., 2010). -->
<!-- All systematic reviews are recommended to be reported following the PRISMA guideline (Page et al, 2021). -->
<!-- All protocols for systematic reviews are recommended to be reported following the PRISMA-P guideline (Page et al, 2021). -->

<!-- In particular when a retrospective study is conducted using hospital records, it is recommended that the RECORD extension of the STROBE guideline is considered (Benchimol et al, 2015). -->

# References

- **SAR-2023-017-BH-v01** -- Sensitivity of mortality rates to the imputation of missing socioeconomic data: cohort study
<!-- - Cohen, J. (1988). Statistical power analysis for the behavioral sciences (2nd Ed.). New York: Routledge. -->
- Zarin DA, et al. The ClinicalTrials.gov results database -- update and key issues. N Engl J Med 2011;364:852-60 (<https://doi.org/10.1056/NEJMsa1012065>).
- Gamble C, et al. Guidelines for the Content of Statistical Analysis Plans in Clinical Trials. JAMA. 2017;318(23):2337–2343 (<https://doi.org/10.1001/jama.2017.18556>).
- von Elm E, Altman DG, Egger M, Pocock SJ, Gøtzsche PC, Vandenbroucke JP; STROBE Initiative. The Strengthening the Reporting of Observational Studies in Epidemiology (STROBE) Statement: guidelines for reporting observational studies. Int J Surg. 2014 Dec;12(12):1495-9 (<https://doi.org/10.1016/j.ijsu.2014.07.013>).
<!-- - Benchimol EI, Smeeth L, Guttmann A, Harron K, Moher D, Petersen I, Sørensen HT, von Elm E, Langan SM; RECORD Working Committee. The REporting of studies Conducted using Observational Routinely-collected health Data (RECORD) statement. PLoS Med. 2015 Oct 6;12(10):e1001885 (<https://doi.org/10.1371/journal.pmed.1001885>). -->
<!-- - Schulz K F, Altman D G, Moher D. CONSORT 2010 Statement: updated guidelines for reporting parallel group randomised trials BMJ 2010; 340 :c332 (<https://doi.org/10.1136/bmj.c332>). -->
<!-- - Page M J, McKenzie J E, Bossuyt P M, Boutron I, Hoffmann T C, Mulrow C D et al. The PRISMA 2020 statement: an updated guideline for reporting systematic reviews BMJ 2021; 372 :n71 (<https://doi.org/10.1136/bmj.n71>). -->
<!-- - Moher D, Shamseer L, Clarke M, Ghersi D, Liberati A, Petticrew M, Shekelle P, Stewart LA. Preferred Reporting Items for Systematic Review and Meta-Analysis Protocols (PRISMA-P) 2015 statement. Syst Rev. 2015;4(1):1 (<https://doi.org/10.1186/2046-4053-4-1>). -->

# Appendix

This document was elaborated following recommendations on the structure for Statistical Analysis Plans (Gamble, 2017) for better transparency and clarity.

## Availability

All documents from this consultation were included in the consultant's Portfolio.

<!-- The client has requested that this analysis be kept confidential until a future date, determined by the client. -->
<!-- All documents from this consultation are therefore not published online and only the title and year of the analysis will be included in the consultant's Portfolio. -->
<!-- After the agreed date is reached, the documents will be released. -->

<!-- The client has requested that this analysis be kept confidential. -->
<!-- All documents from this consultation are therefore not published online and only the title and year of the analysis will be included in the consultant's Portfolio. -->

The portfolio is available at:

<https://philsf-biostat.github.io/SAR-2023-017-BH/>

## Associated analyses

This analysis is part of a larger project and is supported by other analyses, linked below.

**Effect of socioeconomic status in mortality rates after brain injury: cohort study**

<https://philsf-biostat.github.io/SAR-2023-004-BH/>

**Time-adjusted effect of socioeconomic status in mortality rates after brain injury: cohort study**

<https://philsf-biostat.github.io/SAR-2023-016-BH/>
